Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-31 tacl.tex
Last fixes before submission

\documentclass[11pt,letterpaper]{article}
\usepackage[letterpaper]{geometry}
\usepackage{acl2012}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{url}
\usepackage{xspace}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}
\makeatletter
\newcommand{\@BIBLABEL}{\@emptybiblabel}
\newcommand{\@emptybiblabel}[1]{}
\makeatother
%\usepackage[hidelinks]{hyperref}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

@@

In this paper we will refer to the targets of our lexicon creation efforts as \termdef{formulaic sequences}, following the terminology of Wray \shortcite{Wray02,Wray08}, wherein a formulaic sequence (FS) is defined as ``a sequence, continuous or discontinuous, of words or other elements, which is, or appears to be, prefabricated: that is, stored and retrieved whole from memory at the time of use, rather than being subject to generation or analysis by the language grammar.'' That is, a FS shows signs of being part of a mental lexicon. Though by this definition individuals or small groups may have their own FS, here we are only interested in FSs that are shared by a recognizable language community.

@@

Our approach to FS identification involves optimization of the total explanatory power of a lattice, where each node corresponds to an $n$-gram type. The explanatory power of the whole lattice is defined simply as a product of the \termdef{explainedness} of the individual nodes. Each node can be considered either ``on'' (is an FS) or ``off'' (is not an FS). The basis of \termdef{explainedness} calculation is the syntax-sensitive LPR association measure of \newcite{Brooke15b}, but it is calculated differently depending on the on/off status of the node as well as the status of the nodes in the vicinity: nodes are linked based on $n$-gram subsumption and corpus overlap relationships (see \figref{fig:example}), with ``on'' nodes typically explaining other nodes.  Given these relationships, we iterate over the nodes and greedily optimize the on/off choice relative to explainedness in the local neighborhood of each node, until convergence.

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. In general, independent FS do not predictably overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together (overlapping on the tokens \ex{keep \gap under}), we do not want both being selected as an FS, even in the case that both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of turned-off nodes, we decrease the explainedness of the overlapping turned-on nodes---a penalty rather than an incentive which expresses the model's confusion at having overlapping FS. Let $\mathit{oc}(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1,\dotts,o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1(t)$, the exponential term for $E_1(t)$, as:
\begin{displaymath}
	d_{1}(t) = \frac{c(t)}{c(t) - \sum_{i=1}^{m}{\mathit{oc}(o_i)}}
\end{displaymath}

@@

Given the relationships between nodes, it is obviously not sufficient to consider switching only the present node. If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of ``relevant'' nodes which are the most important to the status of the node under consideration. Since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes, starting from an `all-on' state and at each iteration turning off the node (if any) which most increases overall explainedness.

@@

       \toprule
			& P & R & F\\
			 \midrule
			PMIrank & 0.20 & 0.29 & 0.23 \\
			minLPRrank & 0.34 & 0.45 & 0.39 \\
			LPR-seg & 0.42 & 0.45 & 0.43 \\
			LocalMaxs & \bf{0.56} & 0.39 & 0.46 \\
			DP-seg & 0.35 & \bf{0.71} & 0.47 \\
  \midrule
			Lattice & 0.47 & 0.63 & \bf{0.54} \\
       \bottomrule
 \end{tabular}
  \caption{ Results of FS identification in contiguous BNC test set; LocalMaxs = method of Silva and Lopes (1999); DP-seg = method of Newman et \al (2012)}
	\label{tab:BNC}

@@

As it stands, connections in the lattice are based entirely on explicit $n$-gram subsumption and overlap relations. One of the major benefits of this model as compared to alternatives is the relative ease with which additional kinds of interactions between nodes could be implemented. Another connection we have considered is based on identical or similar syntactic patterns (\eg POS sequences), which could serve to encourage the model to make generalizations: in English, \eg learning that verb-particle combinations are generally likely to be FS, whereas verb-determiner combinations are not. Our initial investigations suggest, however, it may be difficult to apply this idea without merely amplifying existing undesirable biases in the LPR measure. Bringing in other information such as simple distributional statistics might help the model identify non-compositional semantics, and could, in combination with the existing lattice competition, focus the model on detecting MWEs, which could in turn serve as a reliable basis for generalization when the raw LPR statistics are an unreliable indicator of formulaticity.

>>>>

\documentclass[11pt,letterpaper]{article}
\usepackage[letterpaper]{geometry}
\usepackage{acl2012}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{url}
\usepackage{xspace}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}
\makeatletter
\newcommand{\@BIBLABEL}{\@emptybiblabel}
\newcommand{\@emptybiblabel}[1]{}
\makeatother
\usepackage[hidelinks]{hyperref}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

@@

In this paper we will refer to the targets of our lexicon creation efforts as \termdef{formulaic sequences}, following the terminology of Wray \shortcite{Wray02,Wray08}, wherein a formulaic sequence (FS) is defined as ``a sequence, continuous or discontinuous, of words or other elements, which is, or appears to be, prefabricated: that is, stored and retrieved whole from memory at the time of use, rather than being subject to generation or analysis by the language grammar.'' That is, a FS shows signs of being part of a mental lexicon. Though by this definition individuals or small groups may have their own FS, here we are only interested in FS that are shared by a recognizable language community.

@@

Our approach to FS identification involves optimization of the total explanatory power of a lattice, where each node corresponds to an $n$-gram type. The explanatory power of the whole lattice is defined simply as a product of the \termdef{explainedness} of the individual nodes. Each node can be considered either ``on'' (is an FS) or ``off'' (is not an FS). The basis of the calculation of explainedness is the syntax-sensitive LPR association measure of \newcite{Brooke15b}, but it is calculated differently depending on the on/off status of the node as well as the status of the nodes in the vicinity: nodes are linked based on $n$-gram subsumption and corpus overlap relationships (see \figref{fig:example}), with ``on'' nodes typically explaining other nodes.  Given these relationships, we iterate over the nodes and greedily optimize the on/off choice relative to explainedness in the local neighborhood of each node, until convergence.

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. In general, independent FS do not consistently overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together (overlapping on the tokens \ex{keep \gap under}), we do not want both being selected as an FS, even in the case that both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of turned-off nodes, we decrease the explainedness of the overlapping turned-on nodes---a penalty rather than an incentive which expresses the model's confusion at having overlapping FS. Let $\mathit{oc}(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1,\dotts,o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1(t)$, the exponential term for $E_1(t)$, as:
\begin{displaymath}
	d_{1}(t) = \frac{c(t)}{c(t) - \sum_{i=1}^{m}{\mathit{oc}(o_i)}}
\end{displaymath}

@@

Given the relationships between nodes, it is obviously not sufficient to consider switching only the present node. If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of ``relevant'' nodes which are the most important to the status of the node under consideration. Since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes, starting from an `all-on' state and at each step turning off the node (if any) which most increases overall explainedness.

@@

       \toprule
			& P & R & F\\
			 \midrule
			PMIrank & 0.20 & 0.29 & 0.23 \\
			minLPRrank & 0.34 & 0.45 & 0.39 \\
			LPR-seg & 0.42 & 0.45 & 0.43 \\
			LocalMaxs & \bf{0.56} & 0.39 & 0.46 \\
			DP-seg & 0.35 & \bf{0.71} & 0.47 \\
  \midrule
			Lattice & 0.47 & 0.63 & \bf{0.54} \\
       \bottomrule
 \end{tabular}
  \caption{ Results of FS identification in contiguous BNC test set; LocalMaxs = method of da Silva and Lopes (1999); DP-seg = method of Newman et \al (2012)}
	\label{tab:BNC}

@@

As it stands, connections in the lattice are based entirely on explicit $n$-gram subsumption and overlap relations. One of the major benefits of this model as compared to alternatives is the relative ease with which additional kinds of interactions between nodes could be implemented. Another connection we have considered is based on identical or similar syntactic patterns (\eg POS sequences), which could serve to encourage the model to make generalizations: in English, \eg learning that verb-particle combinations are generally likely to be FS, whereas verb-determiner combinations are not. Our initial investigations suggest, however, it may be difficult to apply this idea without merely amplifying existing undesirable biases in the LPR measure. Bringing in other information such as simple distributional statistics might help the model identify non-compositional semantics, and could, in combination with the existing lattice competition, focus the model on detecting MWEs, which could in turn serve as a reliable basis for generalization when the raw LPR statistics alone are an unreliable indicator of formulaticity.
